Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-06-29 metrics.tex
Making corrections..

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total processing time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
R_{proc} = \frac{gB}{T_{proc}} ~ [\mathrm{Bytes/second}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represent the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in \cite{soerensen2016sparse}.
The purpose of the benchmarks is to quantify the processing time for
any of the coding schemes considered in Section \ref{sec:schemes}.

\subsubsection{Encoding Benchmark}
Fig. \ref{fig:enc_goodput_benchmark} refers to the benchmark setup
made for measuring the encoding processing time. The time benchmark
is divided in two parts. In the first part, called a \textit{pre-run},
we quantify the amount of \textit{transmitted} coded packets from an
encoder (named 1 in Fig. \ref{fig:enc_goodput_benchmark}) required for
decoding in a single encoder-decoder link with no packet erasures
for a defined configuration of coding scheme, coding parameters and
amount of feedback. In the case of \ac{TSNC}, we also record the
points where a feedback was received from the decoder. The purpose of
the pre-run is to predict how an encoder should behave with a given
seed within its random number generator. This part of the benchmark is
not regarded as a measurement.

In the second part which is the actual \textit{simulation}, a reseted
encoder (2 in Fig. \ref{fig:enc_goodput_benchmark}) is given the same
seed as in the pre-run and then we measure the time from which the
process start until we reach the amount of transmitted packets in the
pre-run. Then, this measurement is stored as the encoder $T_{proc}$
for this generic configuration.

@@

In the pre-run, we still quantify the amount of transmitted coded
packets from encoder 1. Notice that we include the feedback case because
it is necesary for the \ac{TSNC} scheme. However, now we store
the \textit{transmitted packets} instead. The reason being
that we want to feed the decoder with the same packets in the same
order. Later, in the actual simulation, an encoder (2 in Fig.
\ref{fig:dec_goodput_benchmark}) is given the same packets in the same
order from the pre-run and then we measure the time from which the
process start until the decoder finishes to retrieve the original
packets. Finally, this measurement is saved as the decoder $T_{proc}$
for this general configuration.

@@

The average power specifies the rate at which energy is
consumed in the \ac{Raspi}. Thus, for a given energy value in the device
battery without any external supplies, this metric permits to infer the amount
of time for which the \ac{Raspi} can operate autonomously before draining
out its battery. For the energy per bit consumption, it indicates how
much energy is expended to effectively transmit or receive 1 bit of
information taking into account the encoding or decoding operations
respectively.

For our energy measurement campaign, we automate the setup presented
in Fig.~\ref{fig:measurement_setup} to sequentially run a series of
simulations for a given configuration of a coding scheme and its parameters,
to estimate the energic expenditure in both of our \ac{Raspi}
models. The energy measurement setup goal is to quantify the power
consumption of the \ac{Raspi} over long periods of processing time to
obtain accurate results. A representative sketch of the setup is shown
in the computer monitor in Fig.~\ref{fig:measurement_setup}.

The energy measurement setup presents a \ac{Raspi} device whose power
supply is an Agilent 66319D \ac{DC} source, instead of a conventional
power chord supply.   To compute the power, we just need to measure
the current since the \ac{Raspi} feeds from a fixed input voltage of $5V$
set by the Agilent device, but its current consumption is variable.
Hence, the measured variable is the current consumed by the device for
this fixed input voltage. The output of the measurements
are later sent to a workstation where the raw measurement data is processed.

@@

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in each the idle state $I_{idle,avg}$
and the processing state $I_{idle,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Amperes}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watts}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we simply express the energy
as the product of the average power by the processing time $T_{proc} [sec/byte]$
obtained from the goodput measurement for the same configuration. In this way, we can
relate the processing time with the goodput, the packet and the generation size
as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joules}]
\end{align}
%

>>>>

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total processing time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
R_{proc} = \frac{gB}{T_{proc}} ~ [\mathrm{Byte/second}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represent the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in \cite{soerensen2016sparse}.
The purpose of the benchmarks is to quantify the processing time for
any of the coding schemes considered in Section \ref{sec:schemes}.

\subsubsection{Encoding Benchmark}
Fig. \ref{fig:enc_goodput_benchmark} refers to the benchmark setup
made for measuring the encoding processing time. The time benchmark
is divided in two parts. In the first part, called a \textit{pre-run},
we quantify the amount of \textit{transmitted} coded packets, from
encoder 1 in Fig. \ref{fig:enc_goodput_benchmark}, required for
decoding in a single encoder-decoder link with no packet erasures
for a defined configuration of coding scheme, coding parameters and
amount of feedback. In the case of \ac{TSNC}, we also record the
points where a feedback was received from the decoder. The purpose of
the pre-run is to predict how an encoder should behave with a given
seed within its random number generator. This part of the benchmark is
not regarded as a measurement.

In the second part which is the actual \textit{simulation}, a reseted
encoder 2 in Fig. \ref{fig:enc_goodput_benchmark} is given the same
seed as in the pre-run and then we measure the time from which the
process start until we reach the amount of transmitted packets in the
pre-run. Then, this measurement is stored as the encoder $T_{proc}$
for this generic configuration.

@@

In the pre-run, we still quantify the amount of transmitted coded
packets from encoder 1. Notice that we include the feedback case because
it is necesary for the \ac{TSNC} scheme. However, now we store
the \textit{transmitted packets} instead. The reason being
that we want to feed the decoder with the same packets in the same
order. Later, in the actual simulation, a new decoder is given the
same packets in the same order from the pre-run and then we measure
the time from which the process starts until the decoder finishes to
retrieve the original packets. Finally, this measurement is saved as
the decoder $T_{proc}$ for this general configuration.

@@

The average power specifies the rate at which energy is
consumed in the \ac{Raspi}. Thus, for a given energy value in the device
battery without any external supplies, this metric permits to infer the amount
of time for which the \ac{Raspi} can operate autonomously before draining
out its battery. For the energy per bit consumption, it indicates how
much energy is expended to effectively transmit or receive one bit of
information taking into account the encoding or decoding operations
respectively.

For our energy measurement campaign, we automate the setup presented
in Fig.~\ref{fig:measurement_setup} to sequentially run a series of
simulations for a given configuration of a coding scheme and its parameters,
to estimate the energic expenditure in both of our \ac{Raspi}
models. The energy measurement setup goal is to quantify the energy
consumption of the \ac{Raspi} over long periods of processing time to
obtain accurate results. A representative sketch of the setup is shown
in the computer monitor in Fig.~\ref{fig:measurement_setup}.

The energy measurement setup presents a \ac{Raspi} device whose power
supply is an Agilent 66319D \ac{DC} source, instead of a conventional
power chord supply.   To compute the power, we just need to measure
the current since the \ac{Raspi} feeds from a fixed input voltage of $5V$
set by the Agilent device, but its electric current demand is variable.
Hence, the measured variable is the current consumed by the device for
this fixed input voltage. The output of the measurements
are later sent to a workstation where the raw measurement data is processed.

@@

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in each the idle state $I_{idle,avg}$
and the processing state $I_{idle,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Ampere}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watt}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we simply express the energy
as the product of the average power by the processing time $T_{proc}
~[\mathrm{sec/byte}]$ obtained from the goodput measurement for the same
configuration. In this way, we can relate the processing time with the
goodput, the packet and the generation size as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joule}]
\end{align}
%
